
The following 2 chunk loads initial data and functions that was provided on:
https://github.com/kendyteng/OpenAccess/tree/main/lifetable_dog_vc_2016
```{r}

if (!require(tidyverse)) { 
  install.packages("tidyverse")
  library(tidyverse)
}

### 1) The simple life table. It can set 'top'

lifetable_0 <- 
  function(data, top = NULL) {
    
    # with no 'top' specified
    if (is.null(top)) {
      D.fraction_life <- data %>%
        #The column used to estimate life expectancy has to be called 'lifespan'
        dplyr::select(lifespan) %>%
        mutate(lifespan_int = floor(data$lifespan),
               fraction_life = lifespan - lifespan_int) %>%
        group_by(lifespan_int) %>%
        dplyr::summarise(ax = mean(fraction_life),
                         sd = sd(fraction_life)) 
      
      D.result <- data.frame(age = D.fraction_life$lifespan_int) %>%
        # dx: number of dogs die in this interval
        mutate(dx = c(table(floor(data$lifespan)))) %>% 
        arrange(-age) %>%
        # lx: number living at this age; qx: proportion of death in this period
        mutate(lx = cumsum(dx),
               qx = dx/lx) %>%
        # ax: fraction of year lived
        left_join(., D.fraction_life, by = c("age" = "lifespan_int")) %>%
        # Lx: number of year lived in the interval
        mutate(Lx = (lx-dx)+ax*dx, 
               # Tx: total number of year lived beyong age x
               Tx = cumsum(Lx),
               # Ex: expected life at age x
               Ex = Tx/lx) %>%
        arrange(age) 
    } 
    
    # With 'top' specified 
    else {
      # Deal with 'top'
      data1 <- data %>%
        mutate(lifespan_int = floor(data$lifespan),
               lifespan_int = if_else(lifespan_int > top, top, lifespan_int))
      
      D.fraction_life <- data1 %>%
        dplyr::select(lifespan, lifespan_int) %>%
        mutate(fraction_life = lifespan - lifespan_int) %>%
        group_by(lifespan_int) %>%
        dplyr::summarise(ax = mean(fraction_life),
                         sd = sd(fraction_life))
      
      D.result <- data.frame(age = D.fraction_life$lifespan_int) %>%
        # dx: number of dogs die in this interval
        mutate(dx = c(table(floor(data1$lifespan_int)))) %>% 
        arrange(-age) %>%
        # lx: number living at this age; qx: proportion of death in this period
        mutate(lx = cumsum(dx),
               qx = dx/lx) %>%
        # ax: fraction of year lived
        left_join(., D.fraction_life, by = c("age" = "lifespan_int")) %>%
        # Lx: number of year lived in the interval
        mutate(Lx = (lx-dx)+ax*dx) %>%
        # Tx: total number of year lived beyong age x
        mutate(Tx = cumsum(Lx)) %>%
        # Ex: expected life at age x
        mutate(Ex = Tx/lx) %>%
        arrange(age) 
    }
    return(D.result)
  }


### 2) The simple life table with 'top', 'cell_min' and 'top_min' being able to be specified

lifetable <- 
  function(data, top = NULL, cell_min = 3, top_min = 11) {
    
    ## Original life table
    the_lt <- lifetable_0(data, top = top)
    
    ## to make sure the the number of observation in  the last year >= 'top_min'
    while (tail(the_lt$dx, 1) < top_min) {
      the_lt <- lifetable_0(data, top = tail(the_lt$age, 1)-1)
    }
    
    ## for 'cell_min'
    if (the_lt$age[1] != 0 | sum(the_lt$dx < cell_min) != 0) {
      reject <- 
        paste0("AT lease one year had the number of obsevarions less than ", cell_min)
      stop(reject)
    }
    return(the_lt)
  }


### 3) The life table with the confidence interval

lifetable_ci <- 
  function(data, top = NULL, size, times, seed = NULL, ci = 0.95, cell_min = 3, top_min = 11) {
    
    the_lt <- lifetable(data, top = top, cell_min = cell_min, top_min =  top_min)
    the_lt1 <- the_lt %>% select(age, the_Ex = Ex) 
    
    set.seed(seed)
    
    ## Iterations
    for (i in seq(times)) {
      d1 <- data %>%
        sample_n(., size = size, replace = T) %>%
        # make sure having the same bottom
        lifetable_0(., top = tail(the_lt$age, 1)) %>%
        select(age, dx, Ex) 
      # take the life table only if all year interval have 'cell_min' or more observations
      if (d1$age[1] == 0 & sum(d1$dx < cell_min) == 0) {
        names(d1)[3] <- str_c("Ex", i)
        the_lt1 <- left_join(the_lt1, d1[, c(1, 3)], by = "age")
      }
    }
    
    ll <- (1-ci)/2
    ul <- 1-ll
    
    ## The confidence interval
    the_lt1 <- the_lt1 %>% 
      # remove columns with NA
      select_if(~ !any(is.na(.))) %>%
      mutate_at(vars(starts_with("Ex")), ~.x-the_Ex) %>%
      select(starts_with("Ex")) %>%
      mutate(lb = apply(., 1, function(x) quantile(x, ll, na.rm = T)),
             ub = apply(., 1, function(x) quantile(x, ul, na.rm = T))) %>%
      select(lb, ub) %>%
      mutate(age = the_lt1$age, ex = the_lt1$the_Ex) %>%
      mutate(lci = lb + ex, uci = ub + ex) %>%
      full_join(the_lt, ., by = c("age", "Ex" = "ex"))
    
    return(the_lt1)
  }

```
```{r}
#####################################
#
# R Codes to generate cohort life table
#
#####################################


### Citation: Kendy T. Teng, Dave C. Brodbelt, Camilla Pegram, David B. Church, Dan G. ONeill. 
###           2021. Life tables of annual life expectancy and mortality for dogs in the United Kingdom. 
###           XXX


### Instruction:
# 1) Load the libraries 
# You will need to install them if you haven't 
# install.packages("rio")
# install.packages("tidyverse")
# install.packages("janitor")
library(rio) # for reading the data
library(tidyverse) # need for the functions as well
library(janitor) 

# You will have to set your working directory and put the files in there.

# 2) Load the functions in 'lifetable_func_oa'
#source('lifetable_func_oa.R')


# 3) Load the data and do some descriptive
df <- import("the_dataset.xlsx")
str(df)
df %>% select_if(is.numeric) %>% map(., summary)
df %>% select_if(is.character) %>% map(., tabyl)
# Important! The column used to estimate life expectancy has to be called 'lifespan'

# 4) Simple complete cohort life table
# 4.1) Give it a try
lt <- lifetable(data = df); lt
# 4.2) Set the last year of the life table with 'top'
lt_1 <- lifetable(data = df, top = 19); lt_1

# 4.3) 'top_min' 
#      *Set the minimal number of observations for the last year (default is 11)
#      *If the last year had fewer than 'top_min', its observations will be combined 
#       with the ones in the previous year.
#      *'top_min' will override 'top' if the number of observations in the year set by 'top' 
#       is less than 'top_min'       
lt_2 <- lifetable(data = df, top_min = 200); lt_2
```
```

### TASK 1 ###
```{r}
data =df
library(ggplot2)
library(moments)
library(dplyr)

find_mode <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}
mean(data$lifespan)
colnames(data)
data %>%group_by(sex) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,1)),med = median(lifespan))
data %>%group_by(neuter) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,1)),med = median(lifespan))
breed_vc_data = data %>%group_by(breed_vc) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,0)),med = median(lifespan)) %>% na.omit() 
View(breed_vc_data[!duplicated(breed_vc_data$breed_vc),])
data %>%group_by(pure_cross) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,1)),med = median(lifespan))
View(na.omit(data) %>%group_by(breed_group) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,1)),med = median(lifespan)))
data %>%group_by(insured) %>%  summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,1)),med = median(lifespan))
ggplot(na.omit(data),aes(x = lifespan, color = pure_cross)) + geom_density()
```


### TASK 2 ###

```{r}
#Kaplan Meier
library(survival)
library(KMsurv)
colnames(data)
#fit <- survfit(Surv(lifespan) ~ sex + neuter +pure_cross +breed_group,data,conf.int=0.95, type="kaplan-meier",conf.type="log",conf.lower = "usual")
KM_model = survfit(Surv(lifespan) ~ sex + neuter +pure_cross +breed_group, data, conf.type = "log")
summary(KM_model)
plot(KM_model, main = "KM model with subsetting",xlab = "time",ylab = "S(x)")

```

```{r}
KM_model = survfit(Surv(lifespan) ~ 1, data,conf.int = 0.95, conf.type = "log")
summary(KM_model)
plot(KM_model, main = "KM estimate with 95% confidence intervals",xlab = "time",ylab = "S(x)")
line(KM_model$lower)
line(KM_model$upper)
```


```{r}
## Nelson Aalen 
NA_model <- survfit(Surv(lifespan) ~ 1, data=data, conf.int = 0.95, type = "fh")
summary(NA_model)


##KM and NA
plot(KM_model, main = "KM v.s. NA estimates", xlab
 = "time", ylab = "S(x)", col = "blue")
lines(NA_model, col = "red")
legend("topright", c("KM", "NA"), col = c("blue", "red"), lty = 1)
```


```{r}
## cox regression
data = data %>%  mutate(yearborn = substring(date_birth,first=1,  last = 4))
colnames(data)
```


```{r}
Cox_model <- coxph(Surv(lifespan) ~yearborn + sex + neuter + pure_cross + breed_vc +breed_group +insured, data=data, method ="breslow")
summary(Cox_model)
```
```{r}
Cox_model <- coxph(Surv(lifespan) ~yearborn + sex + neuter + pure_cross  +breed_group +insured, data=data, method ="breslow")
summary(Cox_model)
```


```{r}

library(survival)
library(survminer)
temp1 = coxph(Surv(lifespan) ~ sex , data=data, method ="breslow")
ggsurvplot(survfit(temp1,data), data,conf.int = TRUE, palette = "Dark2", censor = FALSE, surv.median.line = "hv")
```



```{r}
base_data = basehaz(Cox_model, centered = FALSE)
print(base_data)
plot(base_data$time, base_data$hazard, xlab = "t", ylab = "lambda_0(t)", main = "Baseline Hazard Rate", type = "s",col = 3) 

```




##### Task 3 ######
```{r}
## insured and uninsured
top1 = round(quantile(data$lifespan,0.99),1)

lt_ci_ins <- lifetable_ci(data = filter(df, insured == "Insured"), times = 50, seed = 1, 
                             size = filter(df, insured == "Insured") %>% nrow(), top = 18)
write.csv(lt_ci_ins,file = "insuredci.csv")

lt_ci_unins <- lifetable_ci(data = filter(df, insured == "Uninsured"), times = 50, seed = 1, 
                             size = filter(df, insured == "Uninsured") %>% nrow(), top = 18)
write.csv(lt_ci_unins,file= "uninsuredci.csv")
lt_ci_unins
```


```{r}
lt_0_ins <- lifetable_0(data = filter(df, insured == "Insured"))
write.csv(lt_0_ins[,c(1,2,3,4,9)],file = "insured.csv")

lt_0_unins <- lifetable_0(data = filter(df, insured == "Uninsured"))
write.csv(lt_0_unins[,c(1,2,3,4,9)],file = "uninsured.csv")

plot(1-lt_0_ins$qx, main = "Survival probability of insured and uninsured pets ", xlab
 = "Year", ylab = "S(x)", col = "blue",type = "S")
lines(1-lt_0_unins$qx,type = "S" ,col = "red")
lines(1-lt_1$qx,type = "S" ,col = "green")
legend("topright", c("Insured", "Uninsured","All"), col = c("blue", "red","green"), lty = 1)
```


```{r}
plot(1-lt_ci_ins$qx, main = "Survival probability of insured and uninsured pets ", xlab
 = "Year", ylab = "S(x)", col = "blue",type = "S")
lines(1-lt_ci_unins$qx,type = "S" ,col = "red")
legend("topright", c("Insured", "Uninsured"), col = c("blue", "red"), lty = 1)
```
```{r}

lt_ci_ins_simp <- lifetable_0(data = filter(tempdata, insured == "Insured"& yearborn <= 2008))


lt_ci_unins_simp <- lifetable_0(data = filter(tempdata, insured == "Uninsured"& yearborn <= 2008))
```




```{r}
plot(1-lt_ci_ins_simp$qx, main = "Survival probability of insured and uninsured pets ", xlab
 = "Year", ylab = "S(x)", col = "blue",type = "S")
lines(1-lt_ci_unins_simp$qx,type = "S" ,col = "red")
legend("topright", c("Insured", "Uninsured"), col = c("blue", "red"), lty = 1)
```



```{r}
### task 4
datatask4 = data %>% mutate(yearborn = substring(date_birth,first=1,  last = 4)) %>%  filter(yearborn <= 2006 )
datatask4

lt_ci_origin <- lifetable_ci_nolimit(data = df, size = nrow(df), times = 50, seed = 1, ci = 0.95)
lt_ci_origin
lt_ci_0004 <- lifetable_ci_nolimit(data = datatask4 , size = nrow(datatask4), times = 50, seed = 1, ci = 0.95)
lt_ci_0004

task4 =  data %>%  mutate(yearborn = substring(date_birth,first=1,  last = 4)) %>%  group_by(yearborn) %>% summarise(avglife = mean(lifespan), n= n(),var = var(lifespan), skew = skewness(lifespan),mode = find_mode(round(lifespan,0)),med = median(lifespan))
View(task4)
ggplot()
```

```{r}
tempdata = data %>%  mutate(yearborn = substring(date_birth,first=1,  last = 4)) %>% mutate(
  group=case_when(
  yearborn < 2000~ 1,
  yearborn < 2005~ 2,
  yearborn < 2010~ 3,
  yearborn < 2015~4))
tempdata
ggplot(tempdata,aes(x = lifespan, color = yearborn)) + geom_density()+labs(title = "life span distribution of dogs born in each year")
ggplot(data,aes(x = yearborn, color = insured)) + geom_histogram()
ggplot(data, aes(x = as.numeric(yearborn), color = insured)) +
  geom_density() + labs(title = "Year of birth of insured and uninsured dogs") + xlab("Year of birth")+ ylab("Density")
ggplot(na.omit(data), aes(x = breed_group, fill = insured)) +
  geom_bar( position=position_dodge()) + xlab("Breed group")
```


```{r}
##old vs whole
KM_model_old = survfit(Surv(lifespan) ~ 1, data %>%  mutate(yearborn = substring(date_birth,first=1,  last = 4)) %>% filter(yearborn <= 2006),conf.int = 0.95, conf.type = "log")
summary(KM_model)
plot(KM_model, main = "Whole data set versus old death data of dogs born prior to 2006", xlab
 = "time", ylab = "S(x)", col = "blue")
lines(KM_model_old, col = "red")
legend("topright", c("Whole", "2006"), col = c("blue", "red"), lty = 1)

```

